---
title: "191029_ Class09"
author: "Mark Allan Co Jacob"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Learning and applying the commands
Using the following commands:
Input: read,view/head
PCA: prcomp
Cluster: kmeans, hclust
Compare: plot,table, etc

Looking at real data.
```{r}
wisc.df <- read.csv("https://bioboot.github.io/bimm143_S18/class-material/WisconsinCancer.csv")

```

```{r}
read.csv("WisconsinCancer.csv", header=TRUE)
```

Looking at this data, we want to look at the number of patients. Using the nrow() commands, we get `r nrow(wisc.df)` patients.

If we want to see how many malignant and benign samples we're dealing with, we can use the table() commands.
```{r}
table(wisc.df$diagnosis)
```



We want to compile the relevant data only, so we don't have to navigate through extra data.
```{r}
# Making a more compact matrix
wisc.data <- as.matrix(wisc.df[,3:32])

# Setting rows names
row.names(wisc.data) <- wisc.df$id
wisc.data
```


```{r}
diagnosis <- wisc.df$diagnosis
colnames(wisc.data)
length(grep("_mean", colnames(wisc.data), value = TRUE))
```

Checking the column means and standard deviation
```{r}
round(colMeans(wisc.data), 3)
# We use use margin = 2 because we're looking at the columns. (ie. Row = 1, Col =2)
round(apply(wisc.data, 2, sd), 3)
# Since the values from either function yielded different values, we must scale.
```

```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)
wisc.pr
summary(wisc.pr)
plot(wisc.pr)

# How to make a PC1 vs PC2
plot(wisc.pr$x[,1],wisc.pr$x[,2])

# How to color it in
plot(wisc.pr$x[,1],wisc.pr$x[,2],col=diagnosis, pch=15)
```


```{r}
biplot(wisc.pr)
plot(wisc.pr$x[, c(1, 3)], col = (diagnosis), 
     xlab = "PC1", ylab = "PC3")
```
Scree plot
```{r}
plot( wisc.pr$x , xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```



```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)
npc
```

```{r}
plot(wisc.pr$x[,1:2], col=diagnosis)
points(npc[,1], npc[,2], col="blue", pch=16, cex=3)
text(npc[,1], npc[,2], c(1,2), col="white")
```

Since red is for malignant and black is benign.
If a patient was in the red zone, we'd want to follow them up as they're being classified to be in the cluster of malignant tumors. 


















