---
title: "Class_7_R_Functions_and_Packages"
author: "Mark Allan Co Jacob"
date: "10/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R functions revisited


To take functions and environments from URLs, one can use the source() function. Once the line has been processed, the global environment is populated with the stored data and functions.
```{r}
source("http://tinyurl.com/rescale-R")
```

After sourcing, we can recall functions:
```{r}
rescale(c(1,10,5,NA,6))
```

Testing the is.numeric() function and the not symbol (!)
```{r}

is.numeric(c(1,5,NA,6))
```
```{r}
is.numeric(c(1,5,NA,6,"Barry"))
```
```{r}
!is.numeric(c(1,5,NA,6,"Barry"))
```
```{r}
!is.numeric(c(1,5,NA,6))
```


## Creating a custom function:
Write a function to find where there are NA elements in two input vectors

First make some simple input where I know the answer. The function **is.na()** was found online. 
```{r}
x <- c(1,2,NA,3,NA)
y <- c(NA,3,NA,3,4)

# is.na specifies whether each object is NA or not through the output of TRUE or FALSE
is.na(x)
is.na(y)


# which() specifies what positions the argument is located
which(is.na(x))
which(is.na(y))
```

Let's be weird, we want to specify where the positions of NA are the same in either x or y
```{r}
is.na(x) & is.na(y)

# sum() function totals the number of TRUEs I have in my input.
sum(is.na(x) & is.na(y))

#Evidence that TRUE equals 1
sum(c(TRUE, TRUE, FALSE, TRUE))
```

Now let's take our snippets into a function:
```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```

```{r}
both_na(x,y)
```

## **eejit** proofing

Testing other conditions to 
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

both_na(x,y1)
both_na(x,y2)

x2 <- c(NA,NA)

#If entities are not of the same length, the both_na function recycles the shorter length to compensate

x <-  c(NA,NA,NA)
y3 <- c(1, NA, NA, NA, NA, NA, NA)
both_na(x,y3)
```


Using length() to solve the recycling issue of the eejit proofing

```{r}

both_na2 <- function (x,y) {
  if(length(x) != length(y)) {
     stop("Your x and y entities are not of the same length")
  } else sum(is.na(x) & is.na(y))

}


```

Writing a function that determines overall grade from each student
```{r}
x <- c(100,90,70, NA)
y <- c(90,80,70)


grade <- function(x) {
  if( any(is.na(x))) {
    warning("Student is missing a homework.")
  }
  mean(x[-which.min(x)], na.rm = TRUE)
}

grade(x)
```

```{r}

url <- "https://tinyurl.com/gradeinput"

Students <- read.csv(url, row.names = 1)

apply(Students, 1, grade)

```

```{r}
x <- c(100,90,70, NA)
y <- c(90,80,70)


grade <- function(x) {
  if( any(is.na(x))) {
    warning("Student is missing a homework.")
  }
  sum(x[-which.min(x)], na.rm = TRUE)/ length(x)
}

grade(x)
```


```{r}
install.packages("BiocManager")
library(BiocManager)
```

```{r}
install.packages("Rcade")
library(Rcade)

install.packages("datapasta")
library(datapasta)
```

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("Rcade")
```

```{r}
devtools::install_github('RLesur/Rcade')
```









